---
title: "Recipe to Create Tilegrams"
description: |
  A recipe to create a Tilegram. This is where landmasses are converted to the
  same size polygon.
author:
  - name: Luke Heley
    url: https://lajh87.github.io
date: 2017-06-26
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This blog post shows how to create tilegrams using simple triganometry and the optimisation algorithm in the `clue` package.


1. Load required packages

```{r, warning=F, message=F}
p <- c("sp", "rgeos","rworldmap","ggplot2", "plotly", "plyr", "clue") # Load required packages
sapply(p, function(x){if(!require(x,char=T)) install.packages(x); require(x,char=T)}); rm(p)
```


2. Load a map of Africa from the `rworldmap` package

```{r}

## Get Africa Data
data("countriesCoarseLessIslands") #  Load simple map without islands
afr <- countriesCoarseLessIslands[which(!is.na(countriesCoarseLessIslands@data$REGION) &  # Subset for Africa
                                          countriesCoarseLessIslands@data$REGION=="Africa"),]

## Project the map to use an equidistance grid
wgs84 <- '+proj=longlat +datum=WGS84'
wrldequi <- "+init=EPSG:32663"
afr <- spTransform(afr, CRS(wrldequi))

```

3. Create helper functions

```{r}
# Function to convert degrees to radians (trigonemetry)
deg2rad <- function(deg) {(deg * pi) / (180)}

# Function to calculate hexagon side length from a given area
hex_side <- function(area) (3^0.25)*sqrt(2*(area/9)) 

# Function to calculate the area of a hexagon given it size
hex_area <- function(side) ((3*sqrt(3))/2*side)

# Function to draw a hexagon
draw_hex <- function(area=hex_area(1), offset_x = 0, offset_y = 0, id=1, tessellate=F){
  side_length <- hex_side(area)
  A <- sin(deg2rad(30)) * side_length
  B <- sin(deg2rad(60)) * side_length
  C <- side_length

  (x <- c(0, 0, B, 2*B, 2*B, B) + (offset_x*B*2) + ifelse(tessellate == T,  B, 0))
  (y <- c(A+C, A, 0, A, A+C, 2*C) + (offset_y*(A+C)))

  Polygons(list(Polygon(coords = matrix(c(x,y),ncol=2),hole = F)),ID=id)
}

# Function to get area of polygons
getArea <-  function(x) {
  getAreaPolygons = function(x) {
    holes = unlist(lapply(x@Polygons, function(x) x@hole))
    areas = unlist(lapply(x@Polygons, function(x) x@area))
    area = ifelse(holes, -1, 1) * areas
    area
  }
  sum(unlist(lapply(x@polygons, getAreaPolygons)))
}

# Function to get the average area of polygons
getAvgArea <- function(x){
  l <- length(x)
  avgArea <- getArea(x)/l
  return(avgArea)
}

# Function to draw hexagon tiles in a tessellated grid
draw_hexTiles <- function(area, offset_x_start=0, offset_x_end=4, offset_y_start=0, offset_y_end =4, id){
  grid <- expand.grid(offset_x_start:offset_x_end, offset_y_start:offset_y_end)
  grid$tessellate <- grid[,2] %% 2 == 0

  hexes <- SpatialPolygons(lapply(1:nrow(grid), function(i){
    draw_hex(area, offset_x = grid[i,1], offset_y = grid[i,2], id =i, tessellate = grid[i,3])

  }))

  names(grid) <- c("offset_x", "offset_y", "tessellate")

  grid <- data.frame(id = 1:nrow(grid),grid)

  SpatialPolygonsDataFrame(hexes, grid)
}


# Function to create tessellated hexagon tiles over a bounding box
hex_tiles <- function(x, cellsize=NULL){

  if(is.null(cellsize)) cellsize <- getAvgArea(x)

  b <- bbox(x)
  dx <- b["x", "max"] - b["x", "min"]
  dy <- b["y", "max"] - b["y", "min"]

  C <- hex_side(cellsize)
  A <- sin(deg2rad(30)) * C
  B <- sin(deg2rad(60)) * C

  hexAcross <- ceiling(dx/(B*2))
  hexUp <- ceiling(dy/((A+C)))

  offset_x_start <- floor(b["x", "min"]/(B*2))
  offset_y_start <- floor(b["y", "min"]/((A+C)))
  offset_x_end <- offset_x_start + hexAcross
  offset_y_end <- offset_y_start + hexUp

  hex_grid <- draw_hexTiles(cellsize, offset_x_start, offset_x_end, offset_y_start, offset_y_end, i)
  proj4string(hex_grid) <- proj4string(x)
  return(hex_grid)
}
```

5. The function below creates a tilegram in R by implementing the 'Hungarian Algorithm' in the `clue` package


```{r}
makeTilegram <- function(sp,cellsize=NULL){

  tiles <- hex_tiles(sp,cellsize) # Create hexagon tiles
 # sp_b <- rgeos::gBuffer(afr, width = 1.1*hex_side(getAvgArea(sp)))
 # tiles <- tiles[sp_b,]

  pts <- rgeos::gCentroid(sp,byid = T)
  pts <- SpatialPointsDataFrame(pts, data.frame(pt_id = row.names(pts), stringsAsFactors = F))

  tileCentroids <- rgeos::gCentroid(tiles, T)
  tileCentroids <- SpatialPointsDataFrame(tileCentroids, data.frame(id = row.names(tileCentroids),stringsAsFactors = F))

  distance <- rgeos::gDistance(tileCentroids, pts, byid=T)
  tile_pref <- t(apply(distance,1, function(x) rank(x,ties.method ="random")))

  solved <- solve_LSAP(tile_pref, maximum = FALSE)
  solved_cols <- as.numeric(solved)

  newDat <- data.frame(tile_region= row.names(tile_pref), id = as.numeric(colnames(tile_pref)[solved_cols]), stringsAsFactors = F)

  newTiles <- tiles
  newTiles@data <- join(newTiles@data, newDat, by="id")
  newTiles <- newTiles[!is.na(newTiles$tile_region),]

  return(newTiles)

}
```

7. Make a tilegram of africa.

```{r}
tileGram <- makeTilegram(afr)
tileGram_f <- fortify(tileGram)
tileGram_f <- join(tileGram_f, tileGram@data,by="id")

p <- ggplot(tileGram_f) +
  geom_polygon(aes(long, lat, group = group, fill=as.numeric(id), text = tile_region)) +coord_equal() +
  theme(legend.position = "none", panel.background = element_blank(), axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())

ggplotly(width = 500,height=500)
```

8. Plot Normal Map for Comparison

```{r}
afr_f <- fortify(afr)
p <- ggplot(afr_f) +
  geom_polygon(aes(long, lat, group = group, fill=as.numeric(as.factor(id)), text = id)) +coord_equal() +
  theme(legend.position = "none", panel.background = element_blank(), axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())

ggplotly(width = 500,height=500)

```

9. Now the world!

```{r}
wrld <- countriesCoarseLessIslands[which(!is.na(countriesCoarseLessIslands@data$REGION) & countriesCoarseLessIslands@data$REGION != "Antarctica"),]
wrldequi <- "+init=EPSG:32663"
wrld <- spTransform(wrld, CRS(wrldequi))

tileGram <- makeTilegram(wrld,cellsize = getAvgArea(afr))

tileGram_f <- fortify(tileGram)
tileGram_f <- join(tileGram_f, tileGram@data,by="id")

p <- ggplot(tileGram_f) +
  geom_polygon(aes(long, lat, group = group, fill=as.numeric(id), text = tile_region)) +coord_equal() +
  theme(legend.position = "none", panel.background = element_blank(), axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())

ggplotly(width = 500,height=500)

wrld_f <- fortify(wrld)
p <- ggplot(wrld_f) +
  geom_polygon(aes(long, lat, group = group, fill=as.numeric(as.factor(id)), text = id)) +coord_equal() +
  theme(legend.position = "none", panel.background = element_blank(), axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())

ggplotly(width = 500,height=500)
```