---
title: "Sentiment Analysis of Public Accounts Committee Oral Evidence Hearing"
description: |
  A short description of the post.
author:
  - name: Luke Heley
    url: https://lajh87.github.io
date: 06-20-2021
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

# Objective

To explore the utility of sentiment analysis in evaluating qualitative information.

# Fetch and Preprocess Data

I extract a [House of Commons Public Accounts Committee Oral Evidence Hearing 
for the Inquiry into Defence Capabilities and the Equipment Plan (2019-2029)][1] 
and tidy the data so that the dialogue is matched to the speaker and the question.

```{r}
document <- httr::GET("https://committees.parliament.uk/oralevidence/429/html/") %>%
  httr::content() %>% 
  xml2::xml_find_all("//p") %>%
  xml2::xml_text()

meta <- document[1:19] # Can this be detected automatically?
text <- document[20:length(document)]

witnesses <- meta[18] %>%
  stringr::str_split(", | and |: |\\.") %>%
  unlist()

witnesses <- witnesses[2:6]
witnesses <- paste0(witnesses,":")
witnesses <- paste(witnesses, collapse = "|")

members <- meta[7:9] %>% 
  stringr::str_split(", | and |: |\\.|and |; ") %>%
  unlist() %>%
  stringr::str_replace(" \\(Chair\\)", "")

stop_words <- c("", "Members present", "NAO Director", 
                "Director of Parliamentary Relations", 
                "Alternate Treasury Officer of Accounts", 
                "were in attendance", "HM Treasury", "NAO",
                "Comptroller & Auditor General",
                "Liaison Committee member also present")

members <- members[!members %in% stop_words]
members <- paste0(members, ":")
members <- paste(members, collapse = "|")

people <- paste(members, witnesses,"Chair:", sep = "|")
questions <- "^Q[0-9]{3}|^Q[0-9]{2}|^Q[0-9]{1}"

tidy_text <- tibble::tibble(text) %>%
  dplyr::mutate(question = stringr::str_extract(text, questions)) %>%
  dplyr::mutate(person = stringr::str_extract(text, people)) %>%
  dplyr::mutate(person = stringr::str_sub(person, 1, nchar(person)-1)) %>%
  dplyr::mutate(dialogue = stringr::str_remove(text, people)) %>%
  dplyr::mutate(dialogue = stringr::str_remove(dialogue, questions)) %>%
  dplyr::mutate(dialogue = stringr::str_trim(dialogue)) %>%
  tidyr::fill(question, person, .direction = "down") %>%
  dplyr::select(question, person, dialogue) %>%
  dplyr::group_by(question, person) %>%
  tidyr::replace_na(list(question = "Opening"))

committee <- meta[2]
oral <- meta[3]
date <- meta[4]

meta_df <- tibble::tibble(committee, oral, date, members, witnesses)

tidy_document <- list(meta_df, body = tidy_text)
tidy_document
```

# Sentiment Analysis

I use Tyler Rinker's excellent [sentimentr][2] package to calculate text polarity
sentiment at the sentence level. This attempts to take into account valence 
shifters (i.e., negators, amplifiers (intensifiers), de-amplifiers (downtoners), 
and adversative conjunctions) while maintaining speed. By default `sentimentr` 
uses the [Jockers 2017][3] dictionary. 

## Sentiment by Person

The plot below shows the average sentiment by person and the variation in 
sentiment. Gareth Davies said the least, but was positive. The chair 
(Meg Hillier) had the most variation.

```{r}
tidy_document$body %>%
  dplyr::mutate(sentence = sentimentr::get_sentences(dialogue)) %$%
  sentimentr::sentiment_by(dialogue, list(person)) %>%
  plot()

```

## Sentiment by Question

The plot below shows the distribution of sentiment by question. The most 
positive question and response was Q71. The least positive was Q6. This is 
most useful in understanding the difficult questions the MOD faces.

```{r}
tidy_document$body %>%
  dplyr::mutate(sentence = sentimentr::get_sentences(dialogue)) %$%
  sentimentr::sentiment_by(dialogue, list(question)) %>%
  plot()
```

## Emotional Valence Over Time

The plot below shows emotional valence over time. The hearing starts with the 
most difficult questions, before getting position

```{r}
tidy_document$body %>%
  dplyr::mutate(sentence = sentimentr::get_sentences(dialogue)) %$%
  sentimentr::sentiment_by(dialogue, list(person)) %>%
  sentimentr::uncombine() %>%
  plot()
```

## Highlighter Test

The `sentimentr` package contains a highlighter function that produces an HTML
document with positive sentences highlighted green and negative sentences 
highlighted red. This is a useful tool to check the analysis. 

```{r, eval = FALSE}
highlight_test <- tidy_document$body %>%
  dplyr::mutate(sentence = sentimentr::get_sentences(dialogue)) %$%
  sentimentr::sentiment_by(dialogue, list(question, person)) 

attr(highlight_test, "averaging.function") <- sentimentr::average_downweighted_zero
sentimentr::highlight(highlight_test,)
```


# Conclusion

Sentiment analysis has merit when analysing large volumes of data. It enable the
user to see the variation in sentiment on aggregate, in groups and over time. This
can help inform where to pay attention to specific elements of text.

[1]: https://committees.parliament.uk/oralevidence/429/html/ 
[2]: https://github.com/trinker/sentimentr/releases/tag/v2.6.1
[3]: https://github.com/mjockers/syuzhet