---
title: "Extract of UK Frigates and Their Precendents"
description: "This post shows how to access the wikipedia knowledge graph"
author: "Luke Heley"
date: "26 March 2023"
freeze: true
format:
  html:
    toc: true
categories:
 - wikipedia
 - scrape
editor_options:
  chunk_output_type: console
---

# Objective

Extract list of UK Frigates and their precedents.

# Data

We use wikidata to first extract the different types of warship

```{r}
q <- '
SELECT ?ship_type ?ship_typeLabel WHERE {
  SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }
  ?ship_type wdt:P31 wd:Q2235308;
    wdt:P279 wd:Q3114762.
  
  OPTIONAL {  }
}
  '

WikidataQueryServiceR::query_wikidata(q)
  
```

And subsequently different types of frigates operated by the Royal Navy

```{r}
q <- '
SELECT ?ship_class ?ship_classLabel ?service_entry ?followed_by ?followed_byLabel ?operator ?operatorLabel ?image ?service_retirement ?follows ?followsLabel ?service_entryLabel ?service_retirementLabel ?imageLabel WHERE {
  SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }
  ?ship_class wdt:P31 wd:Q559026; # instance of ship class
    wdt:P279 wd:Q106070703. # subclass of frigate
  ?ship_class wdt:P137 wd:Q172771. # operated by RN
  OPTIONAL { ?ship_class wdt:P729 ?service_entry. }
  OPTIONAL { ?ship_class wdt:P730 ?service_retirement. }
  OPTIONAL { ?ship_class wdt:P155 ?follows. }
  OPTIONAL { ?ship_class wdt:P156 ?followed_by. }
  OPTIONAL { ?ship_class wdt:P137 ?operator. }
  OPTIONAL { ?ship_class wdt:P18 ?image. }
}
'
rn_frigates <- WikidataQueryServiceR::query_wikidata(q)
```

```{r}
rn_frigates |>
  dplyr::filter(operatorLabel == "Royal Navy") |>
  dplyr::select(from = ship_classLabel, 
                to = followed_byLabel, 
                value = service_entry) |>
  dplyr::distinct() |>
  dplyr::filter(!is.na(to)) |>
  igraph::graph_from_data_frame() |>
  visNetwork::visIgraph()
```

Next step. Extract data on type 26, type 23, type 22, type 21, leander class, rothesay-class, valour-class, whiteby-class and saisbury class.

Future intention is to compare to other nations where data are available.
