---
title: "Cluster Points based on Distance"
description: |
  A recipe to cluster coordinates based on a specified distance between them.
author:
  - name: Luke Heley
    url: https://lajh87.github.io
date: 03-23-2021
output: distill::distill_article
---

Load [margrittr](https://magrittr.tidyverse.org/). Functions from other packages
are called directly with `::`.

```{r setup}
library(magrittr)
```

Generate function to calculate new coordinates given current coordinates an 
angle in degrees (a) and a given distance (d).

```{r}
#' Convert degrees to radians
deg2rad  <- function(x) {x * pi/180} 

#' Calculate new coordinates for point given angle and distance
distance_from_point <- function(coord, a, d){
  x <- coord[1]
  y <- coord[2]
  
  xx <- x + (d * cos(deg2rad(a)))
  yy <- y + (d * sin(deg2rad(a)))
  
  c(x = xx, y = yy)
} 

```

Generate points with two distinct clusters and plot.

```{r}
label <- c("A", "B", "C", "D", "E", "F", "G")
nodes <- dplyr::bind_rows(
  distance_from_point(c(0,0), 0, 0),
  distance_from_point(c(0,0), 60, -10),
  distance_from_point(c(0,0), 120, -10),
  distance_from_point(c(0,0), 120, -20),
  distance_from_point(c(0,0), 120, -30),
  distance_from_point(c(40,0), 0, 0),
  distance_from_point(c(40,0), 90, -10)
  ) %>%
  dplyr::mutate(id = label) %>%
  dplyr::select(.data$id, .data$x, .data$y)


ggplot2::ggplot(nodes, ggplot2::aes(x, y, label = id)) + 
  ggplot2::geom_point() +
  ggplot2::geom_label()

```

Calculate the distance between each point and plot.

```{r}
links <- tidyr::expand_grid(from = label, to = label) %>%
  dplyr::left_join(
    nodes %>% 
      dplyr::rename(from = .data$id, 
                    from_x = .data$x, 
                    from_y = .data$y),
    by = "from"
  ) %>%
  dplyr::left_join(
    nodes %>% 
      dplyr::rename(to = .data$id, 
                    to_x = .data$x, 
                    to_y = .data$y),
    by = "to"
  ) %>%
  dplyr::mutate(distance = sqrt((.data$to_x - .data$from_x)^2 + 
                                  (.data$to_y - .data$from_y)^2)) 

links10m <- links %>% dplyr::filter(.data$distance <= 10)
ggplot2::ggplot(links10m) + 
  ggplot2::geom_segment(ggplot2::aes(x = from_x, y = from_y, 
                                     xend = to_x, yend = to_y)) + 
  ggplot2::geom_label(data = nodes, ggplot2::aes(x = x, y = y, label = label))

```

Loop through the unique labels and list the points that are within a specified
distance (10 units).

```{r}
groups <- label %>% 
  purrr::map(function(x){
   links10m %>% 
      dplyr::filter(.data$from == x) %>% dplyr::pull(.data$to)
  })

pander::pander(groups)

```

For each of the groups check if any of the points in the group are within any 
of the other groups. If they are then concatinate the vector with `union`. 
Filter the unique variables

```{r}
clusters <- groups %>%
  purrr::map(function(x){
    purrr::map(groups, function(y) if(any(unlist(x) %in% unlist(y))) union(x,y)) %>%
      unlist() %>% unique() %>% stringr::str_sort()
  }) %>%
  unique()

pander::pander(clusters)
```

## Alternative Method 1

Create a graph and perform a depth first search

```{r}

g <- igraph::graph_from_data_frame(links10m, directed = TRUE, vertices = nodes)
plot(g)

purrr::map(label, function(x){ 
  ord <- igraph::dfs(g, x, unreachable = FALSE)$order %>% as.numeric()
  ord <- ord[!is.na(ord)]
  label[ord] %>% stringr::str_sort()
}) %>%
  unique()

```

## Alternative Method 2

Draw a circle with radius equal to distance / 2 and merge overlaping and 
adjacent shapes.

## Benchmarking 

```{r}
system.time({
  label %>% 
  purrr::map(function(x){
   links10m %>% 
      dplyr::filter(.data$from == x) %>% dplyr::pull(.data$to)
  }) %>%
    purrr::map(function(x){
    purrr::map(groups, function(y) if(any(unlist(x) %in% unlist(y))) union(x,y)) %>%
      unlist() %>% unique() %>% stringr::str_sort()
  }) %>%
  unique()
})

system.time({
  g <- igraph::graph_from_data_frame(links10m, directed = TRUE, vertices = nodes)
  purrr::map(label, function(x){ 
    ord <- igraph::dfs(g, x, unreachable = FALSE)$order %>% as.numeric()
    ord <- ord[!is.na(ord)]
    label[ord] %>% stringr::str_sort()
  }) %>%
  unique()
})

```